### 2019-02-04 doc/paper/glossary.tex
draft version

\usepackage{glossaries}

\newacronym{utc}{UTC}{Coordinated Universal Time}
\newacronym{adt}{ADT}{Atlantic Daylight Time}
\newacronym{est}{EST}{Eastern Standard Time}
\newglossaryentry{fido2}{name=FIDO2, description=New standart from the FIDO Alliance which combines CTAP and WebAuthn}
\newacronym{fido}{FIDO}{Fast IDentity Online}
\newacronym{ctap}{CTAP}{Client-to-Authenticator Protocol}
\newacronym{nist}{NIST}{National Institute of Standards and Technology}
\newacronym{webAuthn}{WebAuthn}{Web Authentication}
\newacronym{u2f}{U2F}{Universal 2nd Factor}
\newacronym{rest}{REST}{REpresentional State Transfer}
\newacronym{api}{API}{Application Programming Interface}
\newacronym{json}{JSON}{JavaScript Object Notation}
\newacronym{cose}{COSE}{CBOR Object Signing and Encryption (COSE)}
\newacronym{cbor}{CBOR}{Concise Binary Object Representation}
\newglossaryentry{sha-256}{name=SHA-256, description=Secure Hash Algorithm 2 with 256 Bit Output}
\newacronym{utf-8}{UTF-8}{Universal Transofmation Format}
\newacronym{nfc}{NFC}{Near Field Communication}
\newacronym{usb}{USB}{Universal Serial Bus}
\newacronym{eid}{eID}{electronical IDentity}
\newacronym{hlos}{HLOS}{High Level Operating System}
\newacronym{aroe}{AROE}{Allowed Restricted Operating System}
\newacronym{mitm}{MitM}{Man in the Middle}
\newacronym{}{}{}